---
title: "Homework 4"
author: "Amy Nguyen"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Clustering and dimension reduction for gene expression data

```{r message=FALSE, results='hide'}
library(tidyverse) 
library(ROCR) 
library(ggridges) 
library(dendextend)
```

```{r, warning=FALSE, message=FALSE}
leukemia_data <- read_csv("leukemia_data.csv")
```

#### (a)
```{r}
leukemia_data <- leukemia_data %>%
  mutate(Type = as.factor(Type))
table(leukemia_data$Type)
```
From the table, the leukemia subtype that occurs the least is BCR-ABL with 15 patients.

#### (b)
```{r}
pr.out = prcomp(leukemia_data[,c(-1)], scale=TRUE, center=TRUE)
pr.var = pr.out$sdev^2
pve <- pr.var / sum(pr.var)

par(mfrow=c(1,2))
plot(pve, xlab="Principal Component",
     ylab="PVE ", type='b', cex=0.5)
plot(cumsum(pve), xlab="Principal Component ",
     ylab=" Cumulative PVE ",  type='b', cex=0.5)
```
```{r}
which(cumsum(pve) >= 0.9)[1]
```
At least 201 PCs are needed in order to explain 90% of the total variation in the data.

#### (c)
```{r}
rainbow_colors <- rainbow(7)
plot_colors <- rainbow_colors[leukemia_data$Type]
plot(pr.out$x[,c(1,2)], col = plot_colors, cex = 0.0001)
text(pr.out$x[,c(1,2)], col = plot_colors, labels = leukemia_data$Type, cex = 0.5)
```

The subtype T-ALL is most clearly separated from the other groups along the PC2 axis.
```{r}
pr.out$rotation[, 1] %>% 
  abs %>% sort(decreasing = TRUE) %>% head(6)
```
The top 6 genes with the highest absolute loadings for PC1 are SEMA3F, CCT2, LDHB, COX6C, SNRPD2, and ELK3.

#### (d)
```{r warning=FALSE}
plot(pr.out$x[,c(1,3)], col = plot_colors, cex = 0.0001)
text(pr.out$x[,c(1,3)], col = plot_colors, labels = leukemia_data$Type, cex = 0.5)
```

Based on the plot, PC3 does seem to do a slightly better job at disciminating between leukemia types than PC2. It is clear in this plot that leukemia of the same subtypes tend to have pretty similar gene expression levels. 

#### (e)
```{r warning=FALSE}
leukemia_subset <- leukemia_data %>% 
  filter(leukemia_data$Type == 'T-ALL' | leukemia_data$Type == 'TEL-AML1' |
           leukemia_data$Type == 'Hyperdip50')
set.seed(1)
hc = hclust(dist(leukemia_subset), method = 'complete')

dend1 <- as.dendrogram(hc)
dend1 %>% 
  color_branches(k=3) %>% 
  color_labels(k=3) %>% 
  set("labels_cex", 0.3) %>% 
  set_labels(labels=leukemia_subset$Type) %>% 
  plot(horiz=TRUE, main="Dendrogram of Three Leukemia Subtypes")
  
dend2 <- as.dendrogram(hc)
dend2 %>% 
  color_branches(k=5) %>% 
  color_labels(k=5) %>% 
  set("labels_cex", 0.3) %>% 
  set_labels(labels=leukemia_subset$Type) %>% 
  plot(horiz=TRUE, main="Dendrogram of Five Leukemia Subtypes")
```
